---
title: "k-means"
author: "Jingting Li"
date: "Nov.01, 2015"
output: pdf_document
---

```{r}
data(wine, package="rattle")
data.train<-wine[-1]
fit.km<-kmeans(data.train,3)
library("fpc")
plotcluster(data.train,fit.km$cluster)
```
The clusters don't seem well-separated.
There are no obvious gaps between clusters, especiallly, there is an overlapping area between cluster 1 and cluster 3.
Use randIndex from flexclust to compare these two parititions – one from data set and one from result of clustering method.
```{r}
confuseTable.km<-table(wine$Type,fit.km$cluster)
library("flexclust")
randIndex(confuseTable.km)
```
We can observe that the result is only around 0.33, and it's far away from 1. So the algorithm doesn't work well.
Use scale function for scaling and centering data.
```{r}
data.trainnew<-scale(wine[-1])
fit.kmnew<-kmeans(data.trainnew,3)
plotcluster(data.trainnew,fit.kmnew$cluster)
```
We can see the data is clustered very well, there are no collapse between clusters. 
```{r}
confuseTable.kmnew<-table(wine$Type,fit.kmnew$cluster)
randIndex(confuseTable.kmnew)
```
The result is quite close to 1, so K-Means works well for classifying wine dataset.


```{r}
data("iris")
data.train<-iris[-5]
fit.km<-kmeans(data.train,3)
plotcluster(data.train,fit.km$cluster)
```
There is an overlapping area between cluster 1 and cluster 3.
```{r}
confuseTable.km<-table(iris$Species,fit.km$cluster)
randIndex(confuseTable.km)
```
The result is not very close to 1.
Using scaled data to cluster the iris dataset again.
```{r}
data.trainnew<-scale(iris[-5])
fit.kmnew<-kmeans(data.trainnew,3)
plotcluster(data.trainnew,fit.kmnew$cluster)
```
There is still an overlap between cluster 1 and cluster 3.
```{r}
confuseTable.kmnew<-table(iris$Species,fit.kmnew$cluster)
randIndex(confuseTable.kmnew)
```
The result is even smaller than that before scaling.
So K-Means doesn't work well for classifying iris dataset. And scaling is not helpful.